---
title: "Assignment 4: Building A Function"
author: "Christina Treacy"
date: "May 14, 2026"
format: pdf
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Contract
This function computes a biodiversity index score using Simpson's Index (D), where inputs are the number of individuals of each species(n_i) and the total number of individuals of all species (N), and outputs are values between 0 and 1. This index measures the probability that two individuals randomly selected from a sample belong to the same species, with lower output values associated with lower diversity.

The function is defined by the following equation:
D = sum((n_i * (n_i - 1)) / (N * (N-1)))

First, we will load my function from the R file I saved it to.

```{r}
library(tidyverse)
library(here)
# source the function from a file 
source(here("HW4/biodiv_index_D.R"))
```

Next, we will calculate values for 3 different species within a population.

```{r}

species1 = biodiv_index_D(143, 685)
species2 = biodiv_index_D(n_i=87, N=685)
species3 = biodiv_index_D(n_i=201, N=685)

species1
species2
species3
```
D for species 1 = 0.04333888
D for species 2 = 0.01596875
D for species 3 = 0.08579844

