---
title: "Week4Practice"
format: html
editor: visual
---

```{r}
library(tidyverse)

```

Write the R code to create test data where we are simulating how many bear-visitor encounters are likely, if we know that

-   the number of encounters per day follows a Poisson distribution with a mean of 1 encounter per day

-   there are 1000 visitors

We want to simulate how many encounters they have Generate sample data and plot the distribution of encounters

```{r}
#| echo: false
bearencounters = rpois(1000, lambda = 1)
ggplot(data.frame(bearencounters=bearencounters), aes(bearencounters))+geom_histogram()
```

Whats the maximum number of encounters any visitor had? --\> 4 encounters

On your own with the airquality data set

1)  find the median Wind speed
2)  Sample 10 Wind speed measurements
3)  compute their median

Submit both on Canvas quiz

```{r}

data()

# Lets look at New York Air Quality Measurement
# load into your R environment
data(airquality)
# learn about dataset
help(airquality)
# a good way to see structure of data
str(airquality)
attributes(airquality)

# does Ozone correlate with air temperature
ggplot(airquality, aes(x=Temp, y=Ozone))+geom_point()+ggtitle("Ozone vs Temperature")

# lets say we are testing a new product and we want a sample of 5 ozone measurement that conforms to the distribution in this dataset
ozone_sample = sample(airquality$Ozone, size=5, replace=F)
ozone_sample
# this works but we have some NA values - maybe subset
airquality_clean = na.omit(airquality)

# Issue 2 - we can't link that Ozone with other measurements in the dataset - we need to sample the row number instead of ozone
ozone_sample = sample(1:nrow(airquality_clean), size=5, replace=F)
ozone_sample
airquality_clean[ozone_sample,]
# plot our sample days

ggplot(airquality_clean, aes(x=Temp, y=Ozone))+geom_point()+
  geom_point(data=data.frame(Temp=airquality_clean$Temp[ozone_sample], Ozone=airquality_clean$Ozone[ozone_sample]), color="red", size=3)

##########################

#median wind speed
median(airquality$Wind) #9.7

#Sample 10 Wind speed measurements
wind_sample = sample(airquality$Wind, size=10, replace=F)
wind_sample #  11.5 10.9  9.7  6.9 16.6  4.6 14.9  8.0 11.5  8.6

#compute their median
median(wind_sample) #10.3

```

In a quarto document

-   create a new population of some creature (could be mythical)
-   sample from the population
-   use summary to explore the distribution of your sample

## But I don't want a factor

Sometimes you don't want factors and R thinks something should be How to change back? **as.numeric** makes sense ...but

```{r fishes}

# create vector of possible fish 
possible.butterfly = c("monarch","swallowtail","painted lady","cabbage white","red admiral", "malachite")

# we can use sample to simulate a random recording of catch by fisherman, lets say we pick 20 fish from the net

catch1 = base::sample(possible.butterfly, size=20, replace=T)
# because possible.fish was a factor catch1 will be a factor
catch1

summary(catch1)
# if we want summary to be more useful - make this a factor
catch1 = as.factor(catch1)
```

cabbage white malachite monarch painted lady red admiral swallowtail 3 2 3 5 4 3

```{r}
# to quickly get frequencies of different fish and to plot 
summary(catch1)
plot(catch1, col="blue")


# we can also use summary to explore and return information about the distribution
# mean frequency of a particular type of fish
mean(summary(catch1))

# maximum frequency
max(summary(catch1))

# which fish was most frequently caught
which.max(summary(catch1))

#to get just the name 
names(which.max(summary(catch1)))

numeric_sample = as.numeric(catch1)
sum(numeric_sample)
numeric_sample

```
